### 2021-08-05 Seldin Report.Rmd
Initial- transfer to work computer

---
title: "Seldin Report: Ssec Scores"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Report on Ssec scores from "Seldin Methods"
  a. Find distribution of Ssec values
  b. Match top Ssec values to gene symbols
  c. Figure out threshold/significance test
  d. Functional Enrichment