### 2021-08-05 Tissue Specific Report.Rmd

---
title: "Report for Tissue-Specific Proteins"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Report results of tissue-specific proteins data

```{r, echo = FALSE, message = FALSE}
# Packages
library(tidyverse)

# Data Input
#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
ts_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
protein_liver = read.csv("Liver_WGCNA.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")
ts_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")

############## Calculate R squared for tissue specific and not tissue specific########################
fin.r.squared_brain = read.csv("R_Squared_Brain.csv")
fin.r.squared_brain_sec = read.csv("R_Squared_Brain_Sec.csv")
fin.r.squared_liver = read.csv("R_Squared_Liver.csv")
fin.r.squared_liver_sec = read.csv("R_Squared_Liver_Sec.csv")
fin.r.squared_brain_notsec = read.csv("R_Squared_Brain_NotSec.csv")
fin.r.squared_liver_notsec = read.csv("R_Squared_Liver_NotSec.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
ts_protein_brain = fix.read.csv(ts_protein_brain)

protein_liver = fix.read.csv(protein_liver)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)
ts_protein_liver = fix.read.csv(ts_protein_liver)

fin.r.squared_brain = fix.read.csv(fin.r.squared_brain)
fin.r.squared_brain_sec = fix.read.csv(fin.r.squared_brain_sec)
fin.r.squared_liver = fix.read.csv(fin.r.squared_liver)
fin.r.squared_liver_sec = fix.read.csv(fin.r.squared_liver_sec)
fin.r.squared_brain_notsec = fix.read.csv(fin.r.squared_brain_notsec)
fin.r.squared_liver_notsec = fix.read.csv(fin.r.squared_liver_notsec)

```

```{r, echo = FALSE, message = FALSE}
# table comparing number of genes
ngenes = ncol(protein_brain)
ngenes_sec = ncol(secreted_protein_brain)
ngenes_ts = ncol(ts_protein_brain)

ngenes.l = ncol(protein_liver)
ngenes_sec.l = ncol(secreted_protein_liver)
ngenes_ts.l = ncol(ts_protein_liver)
tot_sec = 1760 # from secreted uniprot data
tot_ts = "WHAT"

num = c(ngenes, ngenes_sec, ngenes_ts, ngenes.l, ngenes_sec.l, ngenes_ts.l, tot_sec, tot_ts)

names(num) = c("Brain: Full Data", "Brain: Secreted Data", "Brain: Tissue-Specific Data", "Liver: Full Data", "Liver: Secreted Data", "Liver: Tissue-Specific Data", "Total Secreted Rat Proteins", "Total Tissue-Specific Rat Proteins")

knitr::kable(num, caption = "Number of Genes", col.names = NA, )

# compare shared genes between secreted brain and secreted liver

gee = intersect(colnames(ts_protein_brain), colnames(ts_protein_liver))

int = c(length(colnames(ts_protein_brain)), length(gee), length(colnames(ts_protein_liver)))
names(int) = c("Tissue-Specific Proteins Brain", "Tissue-Specific Proteins Shared by Brain and Liver", "Tissue-Specific Proteins Liver")

knitr::kable(int)

## THis isn't right...

```