### 2021-08-07 Seldin Report.Rmd
Permuted wrong

---
title: "Seldin Report: Ssec Scores"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Report on Ssec scores from "Seldin Methods"
  a. Find distribution of Ssec values
  b. Match top Ssec values to gene symbols
  c. Figure out threshold/significance test
  d. Functional Enrichment

* Using exclusive ts proteins unless we decide later to change back to Uniprot TS

```{r, echo = FALSE, message = FALSE}

library(tidyverse)
library(reshape2)
library(WGCNA)
#library(reshape)
#library(ggplot2)

# Data Input

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_liver = read.csv("Liver_WGCNA.csv")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")
sec_annot = read.csv("Sec_Annotation.csv")
tissuespec_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
tissuespec_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
ts_protein_brain = read.csv("TissueSpecific2_Protein_Brain.csv")
ts_protein_liver = read.csv("TissueSpecific2_Protein_Liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

protein_liver = fix.read.csv(protein_liver)
protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)
sec_annot = fix.read.csv(sec_annot)
tissuespec_protein_brain = fix.read.csv(tissuespec_protein_brain)
tissuespec_protein_liver = fix.read.csv(tissuespec_protein_liver)
ts_protein_brain = fix.read.csv(ts_protein_brain)
ts_protein_liver = fix.read.csv(ts_protein_liver)

```
Calculating Ssec values
```{r, echo = FALSE, message = FALSE}

#### BRAIN ####

brain.liv = bicorAndPvalue(secreted_protein_brain, protein_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv$p))

Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

## tissue-specific genes

ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
top_brain = head(ts_Ssec_brain[,1])

#### LIVER ####

liv.brain = bicorAndPvalue(secreted_protein_liver, protein_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain$p))

Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

## tissue-specific genes

ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
top_liver = head(ts_Ssec_liver[,1])
bot_liver = tail(ts_Ssec_liver[,1])

```

PROCEED WITH CAUTION
```{r}
####################################
# Calc Ssec for all proteins# Don't do this !!

brain.liv = bicorAndPvalue(protein_brain, protein_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv$p))

Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

## tissue-specific genes

top_brain = head(Ssec_brain[,1])
```

```{r}
# Visualizations
hist(Ssec_brain$Ssec)
hist(ts_Ssec_brain$Ssec)

h1 = cbind(rep("brain", times = length(Ssec_brain$Ssec)), Ssec_brain$Ssec)
h2 = cbind(rep("TS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)

compare = rbind(h1, h2)
colnames(compare) = c("trt", "val")
compare = as.data.frame(compare)
compare$trt = as.factor(compare$trt)
compare$val = as.numeric(compare$val)

ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Ssec Distribution: Secreted Proteins v. TS Proteins", x = "Ssec", y = "Density")

plotting = melt(liv.brain$p, id.vars = "refseq")
limit = plotting[as.character(plotting$Var1) %in% c(top_liver, bot_liver),]

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="P-Value Density Curve: Top6 and Bottom6 Liver Proteins", x = "P-Value", y = "Density")

```
NEXT: permute null distribution of Ssec; Calculate Ssec with cutoff; maybe figure out how to run Ssec on non-secreted proteins; null distribution of p-values?

We permuted the target tissue strains and recalculated Ssec to
construct the null distribution and maintain the gene-gene correlation structures in each tissue. These tests typically required
1,000–1,000,000 permutations, listed as p values in Table S1.
The permutation tests were designed to specifically to allow generation of a statistic while maintaining correlation structure to
limit the influence of inflation of cross-tissue gene correlations
(discussed below). We note that in these tests the p values
used reflect a placeholder for the correlation coefficient, which
does not require independence.

```{r}
### This is no good

brain.liv_p = brain.liv$p

brain.liv_perm = apply(brain.liv_p, 2, sample, size = dim(brain.liv_p)[1], replace = F)
rownames(brain.liv_perm) = as.character(1:dim(brain.liv_perm)[1])

# calc Ssec

scores = rowSums(-log(brain.liv_perm))

Ssec_brain_perm = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

hist(Ssec_brain$Ssec)
hist(Ssec_brain_perm$Ssec)

```
```{r}

#permute strain names in target tissue array
# bicor> pval> Ssec val
# repeat 1000- compare observed value to mean of permuted values

```