### 2021-08-09 Tissue Specific Report.Rmd
final TS report

---
title: "Report for Tissue-Specific Proteins"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Report results of tissue-specific proteins data

The ensembl rat liver and brain RNAseq data was downloaded from BioMart database (accessed 2021). Inbred strains were excluded, leaving only recombinant inbred strains. Enrichment data was retrieved from BiomaRt (version number 2.48.0), which was used to exclude non-protein-coding genes from the data set. This resulted in a dataset of 30 strains and 16,457 genes for brain tissue, and 30 strains and 14,109 genes for liver tissue.

From there, the data was reduced to secreted proteins, then to tissue-specific secreted proteins. Results for protein-coding genes were filtered by species "Rattus norvegicus" and subcellular location "secreted" from the uniprot database (accessed 2021). The uniprot IDs from the new secreted dataset were compared to the converted uniprot IDs (via biomaRt) of the brain and liver tissue datasets. From the secreted proteins datasets, we then determined the proteins in the brain and liver datasets were exclusive to those datasets (only found in the brain dataset OR the liver dataset). This resulted in two tissue-specific secreted protein datasets, with 174 in the brain dataset, and 35 in the liver dataset.

The following report was completed using `r R.Version()$version.string`. Additionally, the packages WGCNA (version 1.70.3), broom (version 0.7.7), tidyverse (version `r packageVersion("tidyverse")`), and knitr (version `r packageVersion("knitr")`).

```{r, echo = FALSE, message = FALSE}
# Packages
library(tidyverse)

# Data Input
#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
#ts_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
protein_liver = read.csv("Liver_WGCNA.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")
#ts_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")

fin.r.squared_brain = read.csv("R_Squared_Brain.csv")
fin.r.squared_liver = read.csv("R_Squared_Liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
#ts_protein_brain = fix.read.csv(ts_protein_brain)

protein_liver = fix.read.csv(protein_liver)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)
#ts_protein_liver = fix.read.csv(ts_protein_liver)

fin.r.squared_brain = fix.read.csv(fin.r.squared_brain)
fin.r.squared_liver = fix.read.csv(fin.r.squared_liver)

```

```{r, echo = FALSE, message = FALSE}
# looking for unique genes

genes_brain = colnames(secreted_protein_brain)
genes_liver = colnames(secreted_protein_liver)

gee = genes_brain %in% genes_liver
gee = gee == FALSE
only_brain = genes_brain[gee]

wee = genes_liver %in% genes_brain
wee = wee == FALSE
only_liver = genes_liver[wee]

ts_protein_brain = secreted_protein_brain[, colnames(secreted_protein_brain) %in% only_brain]
ts_protein_liver = secreted_protein_liver[, colnames(secreted_protein_liver) %in% only_liver]

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
write.csv(ts_protein_brain, "TissueSpecific2_Protein_Brain.csv")
write.csv(ts_protein_liver, "TissueSpecific2_Protein_Liver.csv")

```

```{r, echo = FALSE, message = FALSE}
# table comparing number of genes
ngenes = ncol(protein_brain)
ngenes_sec = ncol(secreted_protein_brain)
ngenes_ts = ncol(ts_protein_brain)

ngenes.l = ncol(protein_liver)
ngenes_sec.l = ncol(secreted_protein_liver)
ngenes_ts.l = ncol(ts_protein_liver)
tot_sec = 1760 # from secreted uniprot data
tot_ts = ngenes_ts + ngenes_ts.l

num = c(ngenes, ngenes_sec, ngenes_ts, ngenes.l, ngenes_sec.l, ngenes_ts.l, tot_sec, tot_ts)

names(num) = c("Brain: Full Data", "Brain: Secreted Data", "Brain: Tissue-Specific Data", "Liver: Full Data", "Liver: Secreted Data", "Liver: Tissue-Specific Data", "Total Secreted Rat Proteins", "Total Tissue-Specific Rat Proteins")

knitr::kable(num, caption = "Number of Genes", col.names = NA, )

# compare shared genes between secreted brain and secreted liver

gee = intersect(colnames(ts_protein_brain), colnames(ts_protein_liver))

int = c(length(colnames(ts_protein_brain)), length(gee), length(colnames(ts_protein_liver)))
names(int) = c("Tissue-Specific Proteins Brain", "Tissue-Specific Proteins Shared by Brain and Liver", "Tissue-Specific Proteins Liver")

knitr::kable(int)

```

```{r, echo = FALSE, message = FALSE}
# R squared

rsq.brain = sub(" ~ strain_factor_brain.*", "", fin.r.squared_brain$names_brain)
rsq.liver = sub(" ~ strain_factor_liver.*", "", fin.r.squared_liver$names_liver)

fin.r.squared_brain[,1] = rsq.brain
fin.r.squared_liver[,1] = rsq.liver

r.squared_brain_ts = fin.r.squared_brain[fin.r.squared_brain$names_brain %in% colnames(ts_protein_brain),]
r.squared_liver_ts = fin.r.squared_liver[fin.r.squared_liver$names_liver %in% colnames(ts_protein_liver),]

r.squared_brain_notts = fin.r.squared_brain[!(fin.r.squared_brain$names_brain %in% colnames(ts_protein_brain)),]
r.squared_liver_notts = fin.r.squared_liver[!(fin.r.squared_liver$names_liver %in% colnames(ts_protein_liver)),]

#brain

gee = cbind(rep("not ts", times = dim(r.squared_brain_notts)[1]), r.squared_brain_notts[,2])
gee2 = cbind(rep("ts", times = dim(r.squared_brain_ts)[1]), r.squared_brain_ts[,2])

wil.b = rbind(gee, gee2)
wil.b = as.data.frame(wil.b)
colnames(wil.b) = c("trt", "r.sq")

wil.b$trt = as.factor(wil.b$trt)
wil.b$r.sq = as.numeric(wil.b$r.sq)

wil.brain = coin::wilcox_test(r.sq ~ trt, wil.b)

# liver

gee = cbind(rep("not ts", times = dim(r.squared_liver_notts)[1]), r.squared_liver_notts[,2])
gee2 = cbind(rep("ts", times = dim(r.squared_liver_ts)[1]), r.squared_liver_ts[,2])

wil.l = rbind(gee, gee2)
wil.l = as.data.frame(wil.l)
colnames(wil.l) = c("trt", "r.sq")

wil.l$trt = as.factor(wil.l$trt)
wil.l$r.sq = as.numeric(wil.l$r.sq)

wil.liver = coin::wilcox_test(r.sq ~ trt, wil.l)

```

The R-squared density curves for the brain data below show similar peaks. The curves for the liver data indicate a greater right skew in the tissue-specific proteins. For both sets of data, the distinction is not as clear as the secreted proteins dataset, likely due to small sample size.

```{r, echo = FALSE, message = FALSE}

# Density R Squared

ggplot(wil.b, aes(x = r.sq, color = trt)) + 
  geom_density() +
  labs(title="R Squared Density Curve: Brain",x="R Squared", y = "Density")

ggplot(wil.l, aes(x = r.sq, color = trt)) + 
  geom_density() +
  labs(title="R Squared Density Curve: Liver",x="R Squared", y = "Density")

# Box R Squared

boxplot(fin.r.squared_brain[,2], r.squared_brain_ts[,2], names = c("Not TS", "TS"), xlab = "R Squared", main = "R Squared Brain Data")
boxplot(fin.r.squared_liver[,2], r.squared_liver_ts[,2], names = c("Not TS", "TS"), xlab = "R Squared", main = "R Squared Liver Data")

```

We tested whether the heritability of the non-tissue-specific and tissue specific datasets were significantly different using the Wilcoxon-Mann-Whitney Test. Both tests showed a significant difference at alpha = 0.05, however the p-value for the brain data was very close to 0.05.

Non-Tissue-Specific Brain Dataset v. Tissue-Specific Brain Dataset:

```{r, echo = FALSE, message = FALSE}
wil.brain

```

Non-Tissue-Specific Liver Dataset v. Tissue-Specific Liver Dataset:

```{r, echo = FALSE, message = FALSE}

wil.liver

```