### 2022-05-23 Intro_Summary.Rmd

# Data Summary of HRDP RNA Data from the Brain and Liver
## Angela Yoder

Goals: Download Phenogen HRDP total RNA ensemble data for Brain and Liver and complete a data summary

```{r, echo = FALSE}
# set working directory
#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

# download brain and liver datasets
brain = read.csv("PhenoGen.HRDP.v5.totalRNA.Brain.gene.ensembl96.txt", header = T, sep = "\t")
liver = read.csv("PhenoGen.HRDP.v5.totalRNA.Liver.gene.ensembl96.txt", header = T, sep = "\t")

# number of samples
numsamples_brain = ncol(brain)
numsamples_liver = ncol(liver)
```

The brain dataset contains `r numsamples_brain` samples. The liver dataset contains `r numsamples_liver` samples.

```{r, echo = FALSE}

# number of strains

  # extracting strain from colname
rats_brain = sub("_.*", "", colnames(brain))
rats_liver = sub("_.*", "", colnames(liver))

  # counting unique strains
numstrains_brain = length(unique(rats_brain))
numstrains_liver = length(unique(rats_liver))

# number of rats in each strain

tbl_brain = table(rats_brain)
tbl_liver = table(rats_liver)

# avg number of rats in each strain

avgsample_brain = mean(tbl_brain)
avgsample_liver = mean(tbl_liver)

ex.avgsample_brain = format(round(avgsample_brain, 2), nsmall = 2)
ex.avgsample_liver = format(round(avgsample_liver, 2), nsmall = 2)

```

There are `r numstrains_brain` strains of rat in the brain dataset, and `r numstrains_liver` in the liver dataset.
There are an average of `r ex.avgsample_brain` samples per strain in the brain dataset, and `r ex.avgsample_liver` in the liver dataset.

```{r, echo = FALSE}

# num genes in each dataset

numgene_brain = nrow(brain)
numgene_liver = nrow(liver)

# number of overlapping genes between datasets

genes_brain = rownames(brain)
genes_liver = rownames(liver)

overlap = match(genes_brain, genes_liver)

overlap = na.omit(overlap)
numoverlap = length(overlap)

ex.numgene_brain = formatC(numgene_brain, big.mark = ",")
ex.numgene_liver = formatC(numgene_liver, big.mark = ",")
ex.numoverlap = formatC(numoverlap, big.mark = ",")

```

There are `r ex.numgene_brain` genes in the brain dataset, and `r ex.numgene_liver` genes in the liver dataset.
There are `r ex.numoverlap` genes that are shared between the two datasets.

```{r, echo = FALSE}

# distribution of gene ranges
range_ay = function(x)
{
  r = range(x)
  m = r[2] - r[1]
  out = 2^m
  return(out)
}

range_brain = apply(brain, 1, range_ay)
range_liver = apply(liver, 1, range_ay)

#range_brain = numeric(dim(brain)[1])
#for (i in 1:dim(brain)[1])
#{
#  r = range(brain[i,])
#  m = r[2] - r[1]
#  range_brain[i] = 2^m
#}

#range_liver = numeric(dim(liver)[1])
#for (i in 1:dim(liver)[1])
#{                                                   
#  r = range(liver[i,])
#  m = r[2] - r[1]
#  range_liver[i] = 2^m
#}

meanrange_liver = mean(range_liver)
meanrange_brain = mean(range_brain)

### testing ###
 #liver[which.max(range_liver), grep('LEW', colnames(liver))]

 outlier_liver = which(range_liver > 32)
 outlier_brain = which(range_brain > 32)

 #find rat names that have outliers
 beep = lapply(brain[outlier_brain,], which.max)
 beepbeep = names(beep)
 #table(beepbeep) All the rats have outliers
 choo = lapply(liver[outlier_liver,], which.max)
 choochoo = names(choo)
 #table(choochoo) All the rats have outliers

 #liver[as.numeric(choo), grep('ACI', colnames(liver))]
 #brain[as.numeric(beep), grep("SS.JrHsd", colnames(brain))]

 #double checking outlier ranges
 max_liver = as.numeric(lapply(liver[outlier_liver,], max))
 min_liver = as.numeric(lapply(liver[outlier_liver,], min))
 checking_liver = max_liver - min_liver

 max_brain = as.numeric(lapply(brain[outlier_brain,], max))
 min_brain = as.numeric(lapply(brain[outlier_brain,], min))
 checking_brain = max_brain - min_brain

# find the largest values for range
sort_range_liver = sort(range_liver)
sort_range_brain = sort(range_brain)

# pull out outliers from liver ranges
outliers = tail(sort_range_liver, n = 3)

# logical vector to weed out outliers
outlier_logical_liver = range_liver > 32
outlier_logical_brain = range_brain > 32

# find indices and names of genes that are outliers
index = c()
name = c()
count = 1
for (i in outlier_logical_liver)
{
  if (i == 0)
  {
    index = c(index, count)
    name = c(name, rownames(liver[count,]))
  }
  count = count + 1
}

# summ is number of values < 5 per bad genes
# geene is the index of the rats that have the values that are too high
new_outlier_liver = liver < 5

#geene = c()
#for (i in index)
#{
#  for (j in 1:length(new_outlier_liver[i,]))
#  {
#    if (new_outlier_liver[i,j] == 0)
#    {
#      geene = c(geene, j)
#    }
#  }
#}

summ = apply(new_outlier_liver[index,], 1, sum)
#apply(new_outlier_liver[index,], 2, grep, pattern = 0)
#apply(new_outlier_liver[index,], 2, )

# max_rat is the maximum value for each rat
#max_rat = c()
#max_index = c()
#numrats = length(liver[1,])
#numgenes = length(liver[,1])
#for (i in 1:numrats)
#{
#  max_rat = c(max_rat, max(liver[,i]))
#  for (j in 1: numgenes)
#  {
#    if (liver[j,i] == max(liver[,i]))
#    {
#      max_index = c(max_index, j)
#    }
#  }
#}

max_rat = apply(liver, 2, max)

#max_gene is the max value for each gene
#max_gene = c()
#for (i in 1: numgenes)
#{
#  max_gene = c(max_gene, max(liver[i,]))
#}

max_gene = apply(liver, 1, max)

log_max_gene = max_gene > 15

# find proportion of range vectors that are outliers
prop_outlier_brain = sum(outlier_logical_brain) - length(outlier_logical_brain)
prop_outlier_liver = sum(outlier_logical_liver) - length(outlier_logical_liver)

# create new vectors with outliers removed with logical vector from before
new_range_liver = outlier_logical_liver * range_liver
new_range_brain = outlier_logical_brain * range_brain

# histograms of new vectors
hist(new_range_brain[new_range_brain != 0], main = "Range of Genes with Outliers Excluded, Brain", xlab = "Range")
hist(new_range_liver[new_range_liver != 0], main = "Range of Genes with Outliers Excluded, Liver", xlab = "Range")
```

outliers greater than 32 (2^5). There were 191 outliers from the brain dataset 228 outliers in the liver dataset

The average range of reads per gene for the brain dataset is `r meanrange_brain` and `r meanrange_liver` for the liver dataset. Above are this histograms for ranges of reads per gene.

```{r, echo = FALSE}

#hist(range_liver, main = "Range of Number of Reads for RNA found in the Liver", xlab = "Number of Reads")
#hist(range_brain, main = "Range of Number of Reads for RNA found in the Brain", xlab = "Number of Reads")

```