### 2022-05-23 EnrichR.Rmd
initial commit figuring out enrichr

---
title: "EnrichR Report"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Determine if Enrichr is a valid replacement for EGSEA for gene set analysis
2. If yes, summarize results of Enrichr

```{r, echo = FALSE, message = FALSE}

# https://cran.r-project.org/web/packages/enrichR/vignettes/enrichR.html

setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")

# top 5 TES genes for brain and liver
# the genes theyre correlated to
# convert to gene names

```

```{r, echo = FALSE}
library(enrichR)

setEnrichrSite("Enrichr") # human genes???
dbs = c("GO_Molecular_Function_2018", "GO_Cellular_Component_2018", "GO_Biological_Process_2018")

enriched = enrichr(sec_Ssec_brain[1:100,], dbs)

enriched

```

MSigDB
https://cran.r-project.org/web/packages/msigdbr/vignettes/msigdbr-intro.html
https://bioinformaticsbreakdown.com/how-to-gsea/