### 2022-05-24 egsea.Rmd

---
title: "EGSEA Analysis"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Use the package EGSEA to find potential biological functions for gene sets
a. Import sets of target genes
b. Run EGSEA analysis

```{r, echo = FALSE}

library(EGSEA)

# load data

setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

ts_Ssec_target_brain = read.csv("Target_TESec_Ssec_brain.csv")
ts_Ssec_target_liver = read.csv("Target_TESec_Ssec_liver.csv")
tsns_Ssec_target_brain = read.csv("Target_TENSec_Ssec_brain.csv")
tsns_Ssec_target_liver = read.csv("Target_TENSec_Ssec_liver.csv")
ts_protein_brain = read.csv("TissueSpecific2_Protein_Brain.csv")
ts_protein_liver = read.csv("TissueSpecific2_Protein_Liver.csv")
tsns_protein_brain = read.csv("NSTissueSpecific_Protein_Brain.csv")
tsns_protein_liver = read.csv("NSTissueSpecific_Protein_Liver.csv")
bicor_brain = read.csv( "Bicor_brain.csv")
bicor_liver = read.csv( "Bicor_liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

ts_Ssec_target_brain = fix.read.csv(ts_Ssec_target_brain)
ts_Ssec_target_liver = fix.read.csv(ts_Ssec_target_liver)
tsns_Ssec_target_brain = fix.read.csv(tsns_Ssec_target_brain)
tsns_Ssec_target_liver = fix.read.csv(tsns_Ssec_target_liver)
ts_protein_brain = fix.read.csv(ts_protein_brain)
ts_protein_liver = fix.read.csv(ts_protein_liver)
tsns_protein_brain = fix.read.csv(tsns_protein_brain)
tsns_protein_liver = fix.read.csv(tsns_protein_liver)
bicor_brain = fix.read.csv(bicor_brain)
bicor_liver = fix.read.csv(bicor_liver)

```

```{r, echo = FALSE}

library(EGSEAdata)

egsea.data("mouse")

```

```{r, echo = FALSE}

# top 4 Ssec values

top_brain = rownames(ts_Ssec_target_brain)[1:5]
top_liver = rownames(ts_Ssec_target_liver)[1:5]

corr_brain = matrix(NA, nrow = 5, ncol = 403)
corr_brain = as.data.frame(corr_brain)

for(i in 1:5)
{
  one = bicor_brain[top_brain[i] == rownames(bicor_brain),]
  corr_brain[i,] = one
}
rownames(corr_brain) = top_brain
colnames(corr_brain) = colnames(bicor_brain)

corr_liver = matrix(NA, nrow = 5, ncol = 2763)
corr_liver = as.data.frame(corr_liver)

for(i in 1:5)
{
  one = bicor_liver[top_liver[i] == rownames(bicor_liver),]
  corr_liver[i,] = one
}

rownames(corr_liver) = top_liver
colnames(corr_liver) = colnames(bicor_liver)

target_forbrain = corr_brain < 0.05
target_forliver = corr_liver < 0.05

num_target_brain = apply(target_forbrain, 1, sum)
num_target_liver = apply(target_forliver, 1, sum)

knitr::kable(num_target_brain, caption = "Number of Liver Target Genes for Each Top Ssec Brain")
knitr::kable(num_target_liver, caption = "Number of Brain Target Genes for Each Top Ssec Liver")

```

WHAT THE FUCK

```{r, echo = FALSE}
# get entrez gene ids

library(biomaRt)

ensembl = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl")

get_attrib <- listAttributes(ensembl)
go_attrib <- get_attrib[grep("GO ",get_attrib$description),]

# getting entrez ids
get_entrez = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype"),
                            #"entrezgene_trans_name",
                            #"entrezgene_id"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl)

get_uniprot = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype",
                            "uniprot_gn_id",
                            "name_1006",
                            "definition_1006","go_linkage_type","namespace_1003"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl)

entrez_tes.brain = get_entrez$entrezgene_id[get_entrez$ensembl_gene_id %in% rownames(ts_Ssec_target_brain)]

sum(get_uniprot$ensembl_gene_id %in% rownames(ts_Ssec_target_brain))
GO_fin = subset(get_uniprot, get_uniprot$ensembl_gene_id %in% rownames(ts_Ssec_target_brain))

entrez_tes.liver = get_entrez$entrezgene_id[get_entrez$ensembl_gene_id %in% rownames(ts_Ssec_target_liver)]

```