### 2022-06-20 EnrichR.Rmd
happy pride month

---
title: "EnrichR Report"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Determine if Enrichr is a valid replacement for EGSEA for gene set analysis
2. If yes, summarize results of Enrichr

```{r, echo = FALSE, message = FALSE}

# https://cran.r-project.org/web/packages/enrichR/vignettes/enrichR.html

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

tesec_brain = read.csv("TE.Sec_Ssec_brain.csv")
tesec_liver = read.csv("TE.Sec_Ssec_liver.csv")
tensec_brain = read.csv("TE.NSec_Ssec_brain.csv")
tensec_liver = read.csv("TE.NSec_Ssec_liver.csv")
bicor_brain = read.csv("Bicor_brain.csv")
bicor_liver = read.csv("Bicor_liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

tesec_brain = fix.read.csv(tesec_brain)
tesec_liver = fix.read.csv(tesec_liver)
tensec_brain = fix.read.csv(tensec_brain)
tensec_liver = fix.read.csv(tensec_liver)
bicor_brain = fix.read.csv(bicor_brain)
bicor_liver = fix.read.csv(bicor_liver)

# top 5 TES genes for brain and liver
# the genes theyre correlated to
# convert to gene names

```

```{r, echo = FALSE}
library(enrichR)

#setEnrichrSite("Enrichr") # human genes???
#dbs = c("GO_Molecular_Function_2018", "GO_Cellular_Component_2018", "GO_Biological_Process_2018")
dbs = c("MSigDB_Computational", "MSigDB_Hallmark_2020", "MSigDB_Oncogenic_Signatures")
fin = c()
for(i in 1:174)
{
  names = colnames(bicor_brain[,bicor_brain[i,] < 0.05])
  enriched = enrichr(tesec_brain[, 1], dbs)
  if(dim(enriched$MSigDB_Computational)[1] != 0 | dim(enriched$MSigDB_Hallmark_2020)[1] != 0 | dim(enriched$MSigDB_Oncogenic_Signatures)[1] != 0 )
  {
    fin = c(fin, i)
  }
}

enriched

```

MSigDB
https://cran.r-project.org/web/packages/msigdbr/vignettes/msigdbr-intro.html
https://bioinformaticsbreakdown.com/how-to-gsea/