### 2022-06-20 MSigDB-SOP-2210-3.Rmd

---
title: "Gene Enrichment Analysis with MSigDB"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1.  Identify rat gene sets
2.  Convert Ensembl Gene IDs into Gene Symbol
3.  Use clusterProfiler to do pathway analysis

https://cran.r-project.org/web/packages/msigdbr/vignettes/msigdbr-intro.html

```{r, echo = FALSE}
library(msigdbr)

all_genes = msigdbr(species = "Rattus norvegicus")

```

```{r, echo = FALSE}

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

tesec_brain = read.csv("TE.Sec_Ssec_brain.csv")
tesec_liver = read.csv("TE.Sec_Ssec_liver.csv")
tensec_brain = read.csv("TE.NSec_Ssec_brain.csv")
tensec_liver = read.csv("TE.NSec_Ssec_liver.csv")
bicor_brain = read.csv("Bicor_brain.csv")
bicor_liver = read.csv("Bicor_liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}
`%!in%` = Negate(`%in%`)

tesec_brain = fix.read.csv(tesec_brain)
tesec_liver = fix.read.csv(tesec_liver)
tensec_brain = fix.read.csv(tensec_brain)
tensec_liver = fix.read.csv(tensec_liver)
bicor_brain = fix.read.csv(bicor_brain)
bicor_liver = fix.read.csv(bicor_liver)

# top 5 TES genes for brain and liver
# the genes theyre correlated to
# convert to gene names

```

Fuck this shit

```{r, echo = FALSE}

library(biomaRt)
ensembl_version = "https://ensembl.org"

ensembl = useMart(biomart = "ENSEMBL_MART_ENSEMBL", dataset = "rnorvegicus_gene_ensembl", version = 109)
ensembl = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl")

get_attrib <- listAttributes(ensembl)
#go_attrib <- get_attrib[grep("GO ",get_attrib$description),]

# getting entrez ids
get_entrez = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype",
                            #"entrezgene_trans_name",
                            "entrezgene_id"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl)

get_uniprot = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype",
                            "uniprot_gn_id",
                            "name_1006",
                            "definition_1006","go_linkage_type","namespace_1003"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl)

entrez_tes.brain = get_entrez[get_entrez$ensembl_gene_id %in% rownames(tesec_brain),]

#sum(get_uniprot$ensembl_gene_id %in% rownames(ts_Ssec_target_brain))
#GO_fin = subset(get_uniprot, get_uniprot$ensembl_gene_id %in% rownames(ts_Ssec_target_brain))

entrez_tes.liver = get_entrez[get_entrez$ensembl_gene_id %in% rownames(tesec_liver),]

# missing gene ids

missing_b = rownames(tesec_brain)[rownames(tesec_brain) %!in% entrez_tes.brain$ensembl_gene_id]

missing_l = rownames(tesec_liver)[rownames(tesec_liver) %!in% entrez_tes.liver$ensembl_gene_id]

```