### 2022-07-01 MSigDB.Rmd
fixing problem with order of tes v. target

---
title: "Gene Enrichment Analysis with MSigDB"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1.  Identify rat gene sets
2.  Convert Ensembl Gene IDs into Entrez Ids
3.  Report on missingness in Biomart
3.  Use clusterProfiler to do pathway analysis

For our gene set enrichment analysis (GSEA), we will be using the data resource MSigDB (Molecular Signatures Database), "a collection of annotated gene sets for use with GSEA software." This dataset was used in conjunction with the GSEA R package clusterProfiler. 
We identified our enrichment dataset via the function msigdbr from the msigdbr r package. The species was identified as "Rattus norvegicus." msigdbr has access to multiple collections, including hallmark gene sets (H), and eight different supplementary gene sets (C1 to C8). In this analysis, we did not exclude any of the collections.

```{r, echo = FALSE}
library(msigdbr)

H_genes = msigdbr(species = "Rattus norvegicus", category = "H")
C2_genes = msigdbr(species = "Rattus norvegicus", category = "C2")
C5_genes = msigdbr(species = "Rattus norvegicus", category = "C5")
C8_genes = msigdbr(species = "Rattus norvegicus", category = "C8")

# https://cran.r-project.org/web/packages/msigdbr/vignettes/msigdbr-intro.html
# https://bioconductor.org/packages/release/bioc/manuals/clusterProfiler/man/clusterProfiler.pdf

```

```{r, echo = FALSE}

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

tesec_brain = read.csv("TE.Sec_Ssec_brain.csv")
tesec_liver = read.csv("TE.Sec_Ssec_liver.csv")
tensec_brain = read.csv("TE.NSec_Ssec_brain.csv")
tensec_liver = read.csv("TE.NSec_Ssec_liver.csv")
bicor_brain = read.csv("Bicor_brain.csv")
bicor_liver = read.csv("Bicor_liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}
`%!in%` = Negate(`%in%`)

tesec_brain = fix.read.csv(tesec_brain)
tesec_liver = fix.read.csv(tesec_liver)
tensec_brain = fix.read.csv(tensec_brain)
tensec_liver = fix.read.csv(tensec_liver)
bicor_brain = fix.read.csv(bicor_brain)
bicor_liver = fix.read.csv(bicor_liver)

# top 5 TES genes for brain and liver
# the genes theyre correlated to
# convert to entrez ids

# top 5 genes for brain and liver

top.brain = tesec_brain[1:5,]
top.liver = tesec_liver[1:5,]

```
We began by identifying the top five brain and liver tissue-exclusive, secreted (TES) protein-coding genes, as identified by Ssec score. We then identified the target tissue transcripts that were most significantly associated with the top TES protein-coding genes (p-value < 0.05). These will be the gene sets for which we will identify annotations. This results in 138 liver genes that are associated with TES protein-coding genes from the brain, and 935 brain genes that are associated with TES protein-coding genes from the liver.

```{r, echo = FALSE}

# genes they are correlated to < 0.05

top.corr.brain = bicor_brain[rownames(bicor_brain) %in% rownames(top.brain),]
top.corr.liver = bicor_liver[rownames(bicor_liver) %in% rownames(top.liver),]

list.oflists = function(mat)
{
  out = list(c(), c(), c(), c(), c())
  for (i in 1:5)
  {
    save = colnames(mat[,mat[i,] < 0.05])
    out[[i]] = save
  }
  return(out)
}

brain.list = list.oflists(top.corr.brain)
liver.list = list.oflists(top.corr.liver)

combine.brain = unique(c(brain.list[[1]], brain.list[[2]], brain.list[[3]], brain.list[[4]], brain.list[[5]]))

combine.liver = unique(c(liver.list[[1]], liver.list[[2]], liver.list[[3]], liver.list[[4]], liver.list[[5]]))

num.brain = c(length(brain.list[[1]]), length(brain.list[[2]]), length(brain.list[[3]]), length(brain.list[[4]]), length(brain.list[[5]]))

num.liver = c(length(liver.list[[1]]), length(liver.list[[2]]), length(liver.list[[3]]), length(liver.list[[4]]), length(liver.list[[5]]))

num.target.brain = data.frame(top.brain[,1], num.brain)
num.target.liver = data.frame(top.liver[,1], num.liver)
colnames(num.target.brain) = c("TES PC from Brain", "Number of Target Genes in Liver")
colnames(num.target.liver) = c("TES PC from Liver", "Number of Target Genes in Brain")

knitr::kable(num.target.brain, caption = "Number of Target Genes in the Liver associated with TES Protein-Coding Genes from the Brain")

knitr::kable(num.target.liver, caption = "Number of Target Genes in the Brain associated with TES Protein-Coding Genes from the Liver")

```

Entrez gene IDs were identified via biomart version 104.

## Missingness Report

When retrieving Entrez IDs for our gene sets of interest, I found that many of the genes within these sets were no longer available within biomaRt. Looking at the Ensembl.org website, many of these ensembl IDs were retired via Ensembl. The version in which these ensembl IDs were available is version 104. This version was used for the enrichment analysis. However, there may be concerns using these retired ensembl IDs, especially where there is no longer a corresponding ID in the new version. Below is a table summarizing the number of genes missing, along with examples of pages from defunct ensembl ID webpages.

```{r, echo = FALSE, eval = FALSE}

library(biomaRt)
#ensembl_version = "https://ensembl.org"

#ensembl = useMart(biomart = "ENSEMBL_MART_ENSEMBL", dataset = "rnorvegicus_gene_ensembl", version = 109)
ensembl.miss = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl")
ensembl = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl", version = 104)

get_attrib <- listAttributes(ensembl)
#go_attrib <- get_attrib[grep("GO ",get_attrib$description),]

# getting entrez ids
get_entrez = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype",
                            #"entrezgene_trans_name",
                            "entrezgene_id"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl.miss)

get_uniprot = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype",
                            "uniprot_gn_id",
                            "name_1006",
                            "definition_1006","go_linkage_type","namespace_1003"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl)

entrez_brain.1= get_entrez[get_entrez$ensembl_gene_id %in% combine.brain,]

#sum(get_uniprot$ensembl_gene_id %in% rownames(ts_Ssec_target_brain))
#GO_fin = subset(get_uniprot, get_uniprot$ensembl_gene_id %in% rownames(ts_Ssec_target_brain))

entrez_liver.1 = get_entrez[get_entrez$ensembl_gene_id %in% combine.liver,]

# missing gene ids

missing_b = combine.brain[combine.brain %!in% entrez_brain.1$ensembl_gene_id]
# 15 missing TES : "ENSRNOG00000047739" "ENSRNOG00000008563" "ENSRNOG00000009140" "ENSRNOG00000038437" "ENSRNOG00000034130" "ENSRNOG00000003887" "ENSRNOG00000007374" "ENSRNOG00000013441" "ENSRNOG00000003891" "ENSRNOG00000038980" "ENSRNOG00000013228" "ENSRNOG00000038150" "ENSRNOG00000049171" "ENSRNOG00000011074" "ENSRNOG00000050172"

# 28 missing target : "ENSRNOG00000003498" "ENSRNOG00000047079" "ENSRNOG00000053527" "ENSRNOG00000057548" "ENSRNOG00000033130" "ENSRNOG00000042543" "ENSRNOG00000046529" "ENSRNOG00000049727" "ENSRNOG00000050499" "ENSRNOG00000050885" "ENSRNOG00000051227" "ENSRNOG00000052192" "ENSRNOG00000054524" "ENSRNOG00000061059" "ENSRNOG00000020805" "ENSRNOG00000003648" "ENSRNOG00000047785" "ENSRNOG00000056194" "ENSRNOG00000013410" "ENSRNOG00000049009" "ENSRNOG00000057089" "ENSRNOG00000059007" "ENSRNOG00000061217" "ENSRNOG00000003054" "ENSRNOG00000042197" "ENSRNOG00000047520" "ENSRNOG00000047821" "ENSRNOG00000048017"

missing_l = combine.liver[combine.liver %!in% entrez_liver.1$ensembl_gene_id]
# 12 missing TES : "ENSRNOG00000020805" "ENSRNOG00000050499" "ENSRNOG00000049727" "ENSRNOG00000033130" "ENSRNOG00000042543" "ENSRNOG00000046529" "ENSRNOG00000013464" "ENSRNOG00000049515" "ENSRNOG00000013410" "ENSRNOG00000016535" "ENSRNOG00000033619" "ENSRNOG00000033245"

# 116 missing target : "ENSRNOG00000038437" "ENSRNOG00000001057" "ENSRNOG00000006591" "ENSRNOG00000007705" "ENSRNOG00000008475" "ENSRNOG00000009241" "ENSRNOG00000019857" "ENSRNOG00000021533" "ENSRNOG00000028576" "ENSRNOG00000038976" "ENSRNOG00000043358" "ENSRNOG00000045720" "ENSRNOG00000045960" "ENSRNOG00000046838" "ENSRNOG00000049878" "ENSRNOG00000052459" "ENSRNOG00000053258" "ENSRNOG00000055404" "ENSRNOG00000056130" "ENSRNOG00000057009" "ENSRNOG00000057250" "ENSRNOG00000057522" "ENSRNOG00000057724" "ENSRNOG00000061051" "ENSRNOG00000009140" "ENSRNOG00000005665" "ENSRNOG00000006226" "ENSRNOG00000007219" "ENSRNOG00000008544" "ENSRNOG00000009835" "ENSRNOG00000010720" "ENSRNOG00000011640" "ENSRNOG00000014847" "ENSRNOG00000024726" "ENSRNOG00000028279" "ENSRNOG00000042321" "ENSRNOG00000042502" "ENSRNOG00000042971" "ENSRNOG00000042985" "ENSRNOG00000043382" "ENSRNOG00000045545" "ENSRNOG00000045767" "ENSRNOG00000045837" "ENSRNOG00000046212" "ENSRNOG00000046272" "ENSRNOG00000047890" "ENSRNOG00000048172" "ENSRNOG00000048322" "ENSRNOG00000048765" "ENSRNOG00000050134" "ENSRNOG00000054121" "ENSRNOG00000055892" "ENSRNOG00000056186" "ENSRNOG00000058047" "ENSRNOG00000058276" "ENSRNOG00000058766" "ENSRNOG00000058984" "ENSRNOG00000059734" "ENSRNOG00000060719" "ENSRNOG00000062048" "ENSRNOG00000002953" "ENSRNOG00000022107" "ENSRNOG00000042211" "ENSRNOG00000047780" "ENSRNOG00000061301" "ENSRNOG00000038150" "ENSRNOG00000038980" "ENSRNOG00000003395" "ENSRNOG00000006051" "ENSRNOG00000009917" "ENSRNOG00000010749" "ENSRNOG00000011856" "ENSRNOG00000012477" "ENSRNOG00000019958" "ENSRNOG00000021183" "ENSRNOG00000022919" "ENSRNOG00000027929" "ENSRNOG00000029837" "ENSRNOG00000031734" "ENSRNOG00000032206" "ENSRNOG00000033894" "ENSRNOG00000034287" "ENSRNOG00000037852" "ENSRNOG00000038184" "ENSRNOG00000039050" "ENSRNOG00000042516" "ENSRNOG00000042939" "ENSRNOG00000046844" "ENSRNOG00000047864" "ENSRNOG00000048722" "ENSRNOG00000049435" "ENSRNOG00000050378" "ENSRNOG00000056555" "ENSRNOG00000057059" "ENSRNOG00000057228" "ENSRNOG00000059539" "ENSRNOG00000060482" "ENSRNOG00000060898" "ENSRNOG00000062013" "ENSRNOG00000000060" "ENSRNOG00000005114" "ENSRNOG00000006320" "ENSRNOG00000009388" "ENSRNOG00000009635" "ENSRNOG00000020821" "ENSRNOG00000021302" "ENSRNOG00000023404" "ENSRNOG00000024678" "ENSRNOG00000046157" "ENSRNOG00000047724" "ENSRNOG00000049802" "ENSRNOG00000050511" "ENSRNOG00000057758" "ENSRNOG00000057852" "ENSRNOG00000058416" "ENSRNOG00000060181"

setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
write.csv(entrez_brain.1, "EntrezIds_brain.csv")
write.csv(entrez_liver.1, "EntrezIds_liver.csv")

```

```{r, echo = FALSE}

missingness = data.frame(dataset = c("TES Brain", "Sig Target Genes Brain", "TES Liver", "Sig Target Genes Liver"), missing = c(15, 28, 12, 116), total = c(174, 138, 35, 930))

knitr::kable(missingness, caption = "Missing Ensembl IDs within Datasets")

```

![Missing Gene: ENSRNOG00000050885](C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data/ENSRNOG00000050885_Ensembl.png)

![Missing Gene: ENSRNOG00000013464](C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data/ENSRNOG00000013464_Ensembl.png)

```{r, echo = FALSE, eval = FALSE}

# duplicates
adding = c()
dup = c()
for (i in entrez_tes.brain$ensembl_gene_id)
{
  if (i %in% adding)
  {
    dup = c(dup, i)
  }
  else
  {
    adding = c(adding, i)
  }

}

dup.entrez = entrez_tes.brain$entrezgene_id[entrez_tes.brain$ensembl_gene_id %in% dup]
# there are 4 NAs in the brain dataset
# there are 3 ensembl duplicates in brain dataset
# there are 5 ensembl duplicates in liver dataset

```

## Enrichment Analysis

```{r, echo = FALSE, message = FALSE, warning = FALSE}

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
entrez_brain = read.csv("EntrezIds_brain.csv")
entrez_liver = read.csv("EntrezIds_liver.csv")

entrez_brain = fix.read.csv(entrez_brain)
entrez_liver = fix.read.csv(entrez_liver)

#msigdbr_t2g = msigdbr_df %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
#enricher(gene = gene_ids_vector, TERM2GENE = msigdbr_t2g, ...)

msigdbr_t2g_H = H_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_C2 = C2_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_C5 = C5_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_C8 = C8_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()

library(clusterProfiler)

# pAdjustMethod, min/max (number?) of genes

#enricher(gene = target_genes, TERM2GENE = msigdbr_t2g, ...)

# H

out.brain_H = list(c(), c(), c(), c(), c())
database = msigdbr_t2g_H

for (i in 1:5)
{
  vect = brain.list[[i]]
  entrez = entrez_brain$entrezgene_id[entrez_brain$ensembl_gene_id %in% vect]
  entrez = na.omit(entrez)
  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
  out.brain_H[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
}

out.liver_H = list(c(), c(), c(), c(), c())

for (i in 1:5)
{
  vect = liver.list[[i]]
  entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
  entrez = na.omit(entrez)
  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
  out.liver_H[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
}

#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
#names(brain.ids) = top.brain[,1]

#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")

#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
#names(liver.ids) = top.liver[,1]

#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")

# C2

out.brain_C2 = list(c(), c(), c(), c(), c())
database = msigdbr_t2g_C2

for (i in 1:5)
{
  vect = brain.list[[i]]
  entrez = entrez_brain$entrezgene_id[entrez_brain$ensembl_gene_id %in% vect]
  entrez = na.omit(entrez)
  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
  out.brain_C2[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
}

out.liver_C2 = list(c(), c(), c(), c(), c())

for (i in 1:5)
{
  vect = liver.list[[i]]
  entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
  entrez = na.omit(entrez)
  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
  out.liver_C2[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
}

#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
#names(brain.ids) = top.brain[,1]

#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")

#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
#names(liver.ids) = top.liver[,1]

#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")

# C5

out.brain_C5 = list(c(), c(), c(), c(), c())
database = msigdbr_t2g_C5

for (i in 1:5)
{
  vect = brain.list[[i]]
  entrez = entrez_brain$entrezgene_id[entrez_brain$ensembl_gene_id %in% vect]
  entrez = na.omit(entrez)
  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
  out.brain_C5[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
}

out.liver_C5 = list(c(), c(), c(), c(), c())

for (i in 1:5)
{
  vect = liver.list[[i]]
  entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
  entrez = na.omit(entrez)
  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
  out.liver_C5[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
}

#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
#names(brain.ids) = top.brain[,1]

#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")

#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
#names(liver.ids) = top.liver[,1]

#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")

# C8

out.brain_C8 = list(c(), c(), c(), c(), c())
database = msigdbr_t2g_C8

for (i in 1:5)
{
  vect = brain.list[[i]]
  entrez = entrez_brain$entrezgene_id[entrez_brain$ensembl_gene_id %in% vect]
  entrez = na.omit(entrez)
  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
  out.brain_C8[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
}

out.liver_C8 = list(c(), c(), c(), c(), c())

for (i in 1:5)
{
  vect = liver.list[[i]]
  entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
  entrez = na.omit(entrez)
  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
  out.liver_C8[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
}

#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
#names(brain.ids) = top.brain[,1]

#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")

#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
#names(liver.ids) = top.liver[,1]

#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")

first.brain = rbind(out.brain_H[[1]], out.brain_C2[[1]], out.brain_C5[[1]], out.brain_C8[[1]])
second.brain = rbind(out.brain_H[[2]], out.brain_C2[[2]], out.brain_C5[[2]], out.brain_C8[[2]])
third.brain = rbind(out.brain_H[[3]], out.brain_C2[[3]], out.brain_C5[[3]], out.brain_C8[[3]])
fourth.brain = rbind(out.brain_H[[4]], out.brain_C2[[4]], out.brain_C5[[4]], out.brain_C8[[4]])
fifth.brain = rbind(out.brain_H[[5]], out.brain_C2[[5]], out.brain_C5[[5]], out.brain_C8[[5]])

first.liver = rbind(out.liver_H[[1]], out.liver_C2[[1]], out.liver_C5[[1]], out.liver_C8[[1]])
second.liver = rbind(out.liver_H[[2]], out.liver_C2[[2]], out.liver_C5[[2]], out.liver_C8[[2]])
third.liver = rbind(out.liver_H[[3]], out.liver_C2[[3]], out.liver_C5[[3]], out.liver_C8[[3]])
fourth.liver = rbind(out.liver_H[[4]], out.liver_C2[[4]], out.liver_C5[[4]], out.liver_C8[[4]])
fifth.liver = rbind(out.liver_H[[5]], out.liver_C2[[5]], out.liver_C5[[5]], out.liver_C8[[5]])

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
#write.csv(first.brain, "Brain Enrichment_1.csv")
#write.csv(second.brain, "Brain Enrichment_2.csv")
#write.csv(third.brain, "Brain Enrichment_3.csv")
#write.csv(fourth.brain, "Brain Enrichment_4.csv")
#write.csv(fifth.brain, "Brain Enrichment_5.csv")

#write.csv(first.liver, "liver Enrichment_1.csv")
#write.csv(second.liver, "liver Enrichment_2.csv")
#write.csv(third.liver, "liver Enrichment_3.csv")
#write.csv(fourth.liver, "liver Enrichment_4.csv")
#write.csv(fifth.liver, "liver Enrichment_5.csv")

```

### Enrichment of Target Liver Protein-Coding Genes acted upon by TES Protein-Coding Genes from Brain

```{r, echo = FALSE}

knitr::kable(first.brain, caption = "Brain TES Protein-Coding Gene: ENSRNOG00000000010")
knitr::kable(second.brain, caption = "Brain TES Protein-Coding Gene: ENSRNOG00000020519")
knitr::kable(third.brain, caption = "Brain TES Protein-Coding Gene: ENSRNOG00000002381")
knitr::kable(fourth.brain, caption = "Brain TES Protein-Coding Gene: ENSRNOG00000016567")
knitr::kable(fifth.brain, caption = "Brain TES Protein-Coding Gene: ENSRNOG00000008176")

```

### Enrichment of Target Brain Protein-Coding Genes acted upon by TES Protein-Coding Genes from Liver

```{r, echo = FALSE}

knitr::kable(first.liver, caption = "Liver TES Protein-Coding Gene: ENSRNOG00000032737")
knitr::kable(second.liver, caption = "Liver TES Protein-Coding Gene: ENSRNOG00000021029")
knitr::kable(third.liver, caption = "Liver TES Protein-Coding Gene: ENSRNOG00000008422")
knitr::kable(fourth.liver, caption = "Liver TES Protein-Coding Gene: ENSRNOG00000002964")
knitr::kable(fifth.liver, caption = "Liver TES Protein-Coding Gene: ENSRNOG00000011973")

```