### 2022-07-01 Update PaCor FU.Rmd

I hate onedrive. I thought I fixed this problem this little shit.

```{r, echo = FALSE}

# Calculating Pearson Correlation

target_liver = cbind(tes.liver, tens.liver)
target_brain = cbind(tes.brain, tens.brain)

pearson.pval = function(ex, why)
{
  test = cor.test(ex, y = why, alternative = "two.sided", method = "pearson")
  return(test$p.value)
}

pear.ssec.brain = c()

for (i in 1:174)
{

  estimate = apply(target_liver, MARGIN = 2, FUN = pearson.pval, why = tes.brain[,i])

  score = sum(-log(estimate))/length(target_liver)

  pear.ssec.brain = c(pear.ssec.brain, score) 
}

pear.ssec.liver = c()

for (i in 1:35)
{

  estimate = apply(target_brain, MARGIN = 2, FUN = pearson.pval, why = tes.liver[,i])

  score = sum(-log(estimate))/length(target_brain)

  pear.ssec.liver = c(pear.ssec.liver, score) 
}

```

```{r, echo = FALSE}

h1 = cbind(rep("Bicor", times = 174), ts_Ssec_target_brain$Ssec)
h2 = cbind(rep("PaCor", times = 174), pacor.ssec.brain.fin$ssec)
h3 = cbind(rep("Pearson", times = 174), pear.ssec.brain)

compare = rbind(h1, h2, h3)
colnames(compare) = c("trt", "val")
compare = as.data.frame(compare)
compare$trt = as.factor(compare$trt)
compare$val = as.numeric(compare$val)

ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="TItle", x = "Ssec", y = "Density")

# liver

h1 = cbind(rep("Bicor", times = 35), ts_Ssec_target_liver$Ssec)
h2 = cbind(rep("PaCor", times = 35), pacor.ssec.liver.fin$ssec)
h3 = cbind(rep("Pearson", times = 35), pear.ssec.liver)

compare = rbind(h1, h2, h3)
colnames(compare) = c("trt", "val")
compare = as.data.frame(compare)
compare$trt = as.factor(compare$trt)
compare$val = as.numeric(compare$val)

ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="TItle", x = "Ssec", y = "Density")

```

```{r, echo = FALSE}
pacor.ssec.brain.sort = pacor.ssec.brain.fin %>%
  arrange(desc(ssec))

pacor.ssec.liver.sort = pacor.ssec.liver.fin %>%
  arrange(desc(ssec))

comparison.liver = data.frame(Bicor_gene = ts_Ssec_target_liver$Gene_symbol, Bicor_Ssec = ts_Ssec_target_liver$Ssec, PaCor_gene = pacor.ssec.liver.sort$gene, Pacor_Ssec = pacor.ssec.liver.sort$ssec)

comparison.brain = data.frame(Bicor_gene = ts_Ssec_target_brain$Gene_symbol, Bicor_Ssec = ts_Ssec_target_brain$Ssec, PaCor_gene = pacor.ssec.brain.sort$gene, Pacor_Ssec = pacor.ssec.brain.sort$ssec)

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

write.csv(comparison.brain, "Compare_Bicor_Pacor_SsecBrain_rank.csv")
write.csv(comparison.liver, "Compare_Bicor_Pacor_SsecLiver_rank.csv")

```

```{r, echo = FALSE, eval = FALSE}

# compare correlation

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

corr.pacorssec.brain = read.csv("Pacor_Brain_TES.csv")
corr.pacorssec.liver = read.csv("Pacor_Liver_TES.csv")

corr.pacorssec.brain = fix.read.csv(corr.pacorssec.brain)
corr.pacorssec.liver = fix.read.csv(corr.pacorssec.liver)

bicor_brain = bicorAndPvalue(tes.brain, target_liver)
bicor_liver = bicorAndPvalue(tes.liver, target_brain)

#######################################################
# CORRELATION
thisone = sample(1:174, 1)
thisone
h1 = cbind(rep("Bicor", times = 403), bicor_brain$bicor[thisone,])
h2 = cbind(rep("PaCor", times = 403), corr.pacorssec.brain[,thisone])

compare = rbind(h1, h2)
colnames(compare) = c("trt", "val")
compare = as.data.frame(compare)
compare$trt = as.factor(compare$trt)
compare$val = as.numeric(compare$val)

ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Correlation Value Bicor v. PaCor", x = "Correlation Value", y = "Density")

# PVALUE
thisone = sample(1:174, 1)
thisone
h1 = cbind(rep("Bicor", times = 403), bicor_brain$p[thisone,])
h2 = cbind(rep("PaCor", times = 403), corr.pacorssec.brain[,thisone])

compare = rbind(h1, h2)
colnames(compare) = c("trt", "val")
compare = as.data.frame(compare)
compare$trt = as.factor(compare$trt)
compare$val = as.numeric(compare$val)

ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Correlation Value Bicor v. PaCor", x = "Correlation Value", y = "Density")

```